---
title: "Preferential sampling with shrimp data"
author: "Fabian E. Bachl"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Set things up

Due to a bug preferential sampling currently only works with the backend-clean branch of inlabru and not using the CRAN version. You can obtain it using
```{r results="hide",warning=FALSE,message=FALSE,eval=FALSE}
devtools::install_github("fbachl/inlabru", ref="backend-clean", dependencies = TRUE)
```
Once you installed inlabru, run
```{r results="hide",warning=FALSE,message=FALSE}
library(INLA)
library(inlabru)
```

Set default INLA values for tutorial session (faster inference but only empirical Bayes):

```{r results="hide",warning=FALSE,message=FALSE}
init.tutorial()
```

## Load the data
Define the spatial domain using a mesh:
```{r warning=FALSE,message=FALSE}
data(shrimp)
ggplot() + gg(shrimp$mesh) + gg(shrimp$hauls) + coord_equal()
```

## Model setup
Integration points for LGCP
```{r warning=FALSE,message=FALSE}
ips = ipoints(shrimp$mesh)
```

Matern model for SPDE covariance
```{r warning=FALSE,message=FALSE}
matern <- inla.spde2.pcmatern(shrimp$mesh, 
                              prior.sigma = c(0.1, 0.01), 
                              prior.range = c(1, 0.01))
```

Latent components
```{r warning=FALSE,message=FALSE}
cmp = ~ spde(map = coordinates, model = matern) + spdeCopy(map = coordinates, copy = "spde", model = matern, fixed = FALSE) + lgcpIntercept + Intercept
```

Likelihoods
```{r warning=FALSE,message=FALSE}
lik1 =  like(components = cmp,
             data =  shrimp$hauls, 
             family = "poisson", 
             formula = catch ~ spde + Intercept)

lik2 =  like(components = cmp,
             data =  shrimp$hauls, 
             family = "cp",
             ips = ips,
             formula = coordinates ~ spdeCopy + lgcpIntercept)

```

## Fit the model
```{r warning=FALSE,message=FALSE}
fit = bru(components = cmp, lik1, lik2, options = list(max.iter = 1))
```

## Results

```{r fig.width=7, message=FALSE, warning=FALSE}
pxl = pixels(shrimp$mesh)
fish.intensity = predict(fit, pxl, ~ exp(spde + Intercept))
ggplot() + gg(fish.intensity) + gg(shrimp$hauls, size = 0.5)
```

